---
title: "Omnivores on a diet. Exploratory analysis for OMNI2"
author: Samuel Coavoux
format: 
  html:
    toc: true
    embed-resources: true
editor: source
---

```{r, include=FALSE}
#| label: setup
library(knitr)
opts_chunk$set(echo = FALSE, 
               warning = FALSE, 
               error = FALSE,
               message = FALSE)

```

```{r packages}
#| label: packages
library(targets)
library(tidyverse)
library(janitor)
library(here)
targets::tar_config_set(store = "../_targets/")
```

```{r}
#| label: load data and recode
# loading data
tar_load("survey", store = "../_targets/")

# make aliases
## For genres names
tar_load("genres_aliases", store = "../_targets/")

## for omnivorousness variables
source(here("R", "common_functions.R"))

# Recode
## Make LCA classes a factor rather than integer/numeric
survey <- survey %>% 
  mutate(across(starts_with("cluster_"), ~paste0("class", str_pad(.x, 2, pad = 0))))

## TEMP: remove NA
survey <- survey %>% 
  filter(cluster_survey != "classNA", 
         cluster_streams != "classNA",
         cluster_streamsprop != "classNA")
```

# First strategy: looking at consumption clusters

## LCA on Surveys

We load the results of the LCA in the survey.

We look at the distribution of classes

```{r}
#| label: cluster from survey, distribution
tabyl(survey$cluster_survey) %>% 
  kable(caption = "Distribution of clusters from LCA on survey data")
```

We interpret the classes based on genre distribution.

```{r}
#| label: plot LCA cluster profiles
#| fig.height: 6
tar_load(gg_lca_profile)
print(gg_lca_profile)
```


We then look at how these classes are distributed in the population, 
across gender, age, and SES.

```{r}
#| label: cluster from survey, demographic data
tar_load(gg_lca_socdem)
gg_lca_socdem
```

Finally, we look at the diversity according to legacy metric/ new sd on legitimacy metric.

```{r}
#| label: cluster from survey, correlation with omnivorousness metrics
#| fig-height: 6
tar_load(gg_lca_omni_paper)
print(gg_lca_omni_paper)
```

## Legitimacy by cluster

Now, how does the legitimacy of artists from some genres vary along clusters?

We look at the distribution of legitimacy genre by genre, cluster by cluster.

TODO: RESCALE THESE BY GENRE (AT THE ARTIST LEVEL) FOR HOMOGENOUS X

```{r}
#| label: legitimacy of specific genres by cluster, survey
#| fig.height: 6
#| fig.width: 8
tar_load(gg_lca_omni_bygenre)
print(gg_lca_omni_bygenre)
```


# Second strategy: looking at continuous metrics


## The link between diversity over legitimacy level and other measures of omnivorousness

```{r}
#| label: pairwise omnivorousness, exo vs. HHI streams
#| fig-height: 3
tar_load(gg_exoomni_by_otheromni)
print(gg_exoomni_by_otheromni)
```

```{r}
#| label: correlations between exo and endo leg

```

## Demographic correlates of omnivorousness

Reminder: HHI omnivorousness is (1-herfindahl index) and indicates the
diversity of music streaming across genres (0 is low, 1 is high).

```{r}
#| label: HHI omnivorousness by gender and degree
tar_load()

select(survey, omni_stream_genres_hhi, gender, degree) %>% 
  filter(!is.na(gender), !is.na(degree)) %>% 
  pivot_longer(-omni_stream_genres_hhi) %>% 
  group_by(name, value) %>% 
  summarize(m = mean(omni_stream_genres_hhi, na.rm=TRUE),
            sd = sd(omni_stream_genres_hhi, na.rm=TRUE),
            n = n(),
            se = 1.96*sd/sqrt(n)) %>% 
  ggplot(aes(value, m, ymin = m-se, ymax=m+se)) +
    geom_point() +
    geom_pointrange() +
    facet_wrap(~name, scales = "free_x") +
    labs(title = "HHI omnivorousness by degree and gender", x="", y = "HHI omnivorousness")
```




## Demographic correlates of exo omnivorousness

```{r}
#| label: exo omnivorousness by gender and degree
select(survey, sd_exo_pca, gender, degree) %>% 
  filter(!is.na(gender), !is.na(degree)) %>% 
  pivot_longer(-sd_exo_pca) %>% 
  group_by(name, value) %>% 
  summarize(m = mean(sd_exo_pca, na.rm=TRUE),
            sd = sd(sd_exo_pca, na.rm=TRUE),
            n = n(),
            se = 1.96*sd/sqrt(n)) %>% 
  ggplot(aes(value, m, ymin = m-se, ymax=m+se)) +
    geom_point() +
    geom_pointrange() +
    facet_wrap(~name, scales = "free_x") +
    labs(title = "Exo omnivorousness by degree and gender", x="", y = "exo omnivorousness")
```





# Misc

Definition of cultural holes omnivorousness

$$EO_i = \sum_{j \in N(i)}(a_{ij} - (\frac{1}{OV_i-1}\sum_{k \in N(i)}^{k \neq j}{o_{jk}}))$$

