---
title: "Omnivores on a diet. Exploratory analysis for OMNI2"
author: Samuel Coavoux
format: 
  html:
    toc: true
    embed-resources: true
editor: source
---

```{r, include=FALSE}
#| label: setup
library(knitr)
opts_chunk$set(echo = FALSE, 
               warning = FALSE, 
               error = FALSE,
               message = FALSE)

```

```{r packages}
#| label: packages
library(targets)
library(tidyverse)
library(janitor)
```

```{r}
#| label: load data and recode
# loading data
tar_load("survey")

# make aliases
## For genres names
tar_load("genres_aliases")

## for omnivorousness variables
source("R/common_functions.R")

# Recode
## Make LCA classes a factor rather than integer/numeric
survey <- survey %>% 
  mutate(across(starts_with("cluster_"), ~paste0("class", str_pad(.x, 2, pad = 0))))

## TEMP: remove NA
survey <- survey %>% 
  filter(cluster_survey != "classNA", 
         cluster_streams != "classNA",
         cluster_streamsprop != "classNA")
```

# First strategy: looking at consumption clusters

## LCA on Surveys

We load the results of the LCA in the survey.

We look at the distribution of classes

```{r}
#| label: cluster from survey, distribution
tabyl(survey$cluster_survey) %>% 
  kable(caption = "Distribution of clusters from LCA on survey data")
```


We first interpret the classes based on the distribution of genres

```{r}
#| label: cluster from survey, interpretation
tb <- survey %>%
  select(hashed_id, cluster_survey, matches("B_genres_\\d+")) %>%
  # consolidate genres
  pivot_longer(matches("B_genres_\\d+")) %>%
  mutate(name = genres_aliases[name],
         value = ifelse(value != "", 1, 0)) %>% 
  # remove empty genres
  filter(!is.na(name)) %>% 
  distinct() %>% 
  group_by(cluster_survey, name) %>% 
  summarise(m=sum(value)/n()) %>% 
  pivot_wider(names_from = cluster_survey, values_from = m) %>% 
  rename(genre = "name")
tb %>% 
  adorn_pct_formatting() %>% 
  kable(caption = "Profils of clusters")
```

```{r}
#| label: plot LCA cluster profiles
#| fig.height: 8
tb %>% 
  pivot_longer(-genre) %>% 
  ggplot(aes(value, genre, fill=genre)) +
    geom_col() +
    facet_wrap(~name)
```

```{r}
#| label: plot LCA cluster profiles most significant genres
#| fig.height: 8

sig_genres <- c("rock", "classical", "jazz", "frenchrap", "frenchsongs")
tb %>% 
  filter(genre %in% sig_genres) %>% 
  pivot_longer(-genre) %>% 
  ggplot(aes(value, genre, fill=genre)) +
    geom_col() +
    facet_wrap(~name)
```

We then look at how these classes are distributed in the population, 
across gender, age, and SES.

```{r}
#| label: cluster from survey, demographic data
cluster_dem <- survey %>% 
  select(cluster_survey, age, gender, degree) %>% 
  filter(!is.na(cluster_survey))
```


```{r}
#| label: cluster from survey, age
cluster_dem %>% 
  ggplot(aes(age, factor(cluster_survey))) +
    geom_boxplot()
```

```{r}
#| label: cluster from survey, degree
cluster_dem %>% 
  filter(!is.na(degree)) %>% 
  ggplot(aes(factor(cluster_survey), fill=degree)) +
    geom_bar(position="fill")
```

```{r}
#| label: cluster from survey, gender
cluster_dem %>% 
  filter(!is.na(gender)) %>% 
  ggplot(aes(factor(cluster_survey), fill=gender)) +
    geom_bar(position="fill")
```

```{r}
#| label: cluster from survey, correlation with omnivorousness metrics
survey %>% 
  select(cluster_survey, starts_with("omni"), mean_exo_pca, sd_exo_pca) %>% 
  pivot_longer(-cluster_survey) %>% 
  mutate(name = recode_vars(name, "legitimacy")) %>% 
  filter(!is.na(value)) %>% 
  group_by(cluster_survey, name) %>% 
  summarize(value = mean(value)) %>% 
  pivot_wider(names_from = name, values_from = value) %>% 
  kable()
```

## Latent Profile Analysis on streaming data 

Let us do the same with the streaming data.

We look at the distribution of classes

```{r}
#| label: cluster from streams, distribution
tabyl(survey$cluster_streams) %>% 
  kable(caption = "Distribution of clusters from LPA on survey data")
```


We first interpret the classes based on the distribution of genres

```{r}
#| label: cluster from streams, interpretation
tb <- survey %>%
  select(hashed_id, cluster_streams, matches("B_genres_\\d+")) %>%
  # consolidate genres
  pivot_longer(matches("B_genres_\\d+")) %>%
  mutate(name = genres_aliases[name],
         value = ifelse(value != "", 1, 0)) %>% 
  # remove empty genres
  filter(!is.na(name)) %>% 
  distinct() %>% 
  group_by(cluster_streams, name) %>% 
  summarise(m=sum(value)/n()) %>% 
  pivot_wider(names_from = cluster_streams, values_from = m) %>% 
  rename(genre = "name")
tb %>% 
  adorn_pct_formatting() %>% 
  kable(caption = "Profils of clusters")
```

```{r}
#| label: plot LPA streams cluster profiles
#| fig.height: 8
tb %>% 
  pivot_longer(-genre) %>% 
  ggplot(aes(value, genre, fill=genre)) +
    geom_col() +
    facet_wrap(~name)
```

```{r}
#| label: plot LPA streams cluster profiles most significant genres
#| fig.height: 8

sig_genres <- c("rock", "classical", "jazz", "frenchrap", "frenchsongs")
tb %>% 
  filter(genre %in% sig_genres) %>% 
  pivot_longer(-genre) %>% 
  ggplot(aes(value, genre, fill=genre)) +
    geom_col() +
    facet_wrap(~name)
```

We then look at how these classes are distributed in the population, 
across gender, age, and SES.

```{r}
#| label: cluster from streams, demographic data
cluster_dem <- survey %>% 
  select(cluster_streams, age, gender, degree) %>% 
  filter(!is.na(cluster_streams))
```


```{r}
#| label: cluster from streams, link with age
cluster_dem %>% 
  ggplot(aes(age, factor(cluster_streams))) +
    geom_boxplot()
```
```{r}
#| label: cluster from streams, link with degree
cluster_dem %>% 
  filter(!is.na(degree)) %>% 
  ggplot(aes(factor(cluster_streams), fill=degree)) +
    geom_bar(position="fill")
```

```{r}
#| label: cluster from streams, link with gender
cluster_dem %>% 
  filter(!is.na(gender)) %>% 
  ggplot(aes(factor(cluster_streams), fill=gender)) +
    geom_bar(position="fill")
```

```{r}
#| label: cluster from streams, link with omnivorousness variables
survey %>% 
  select(cluster_streams, starts_with("omni"), mean_exo_pca, sd_exo_pca) %>% 
  pivot_longer(-cluster_streams) %>% 
  mutate(name = recode_vars(name, "legitimacy")) %>% 
  filter(!is.na(value)) %>% 
  group_by(cluster_streams, name) %>% 
  summarize(value = mean(value)) %>% 
  pivot_wider(names_from = name, values_from = value) %>% 
  kable()
```


## Streaming data -- proportioned

And now with the proportioned streaming data

We look at the distribution of classes

```{r}
#| label: cluster from streams with proportion, distribution
tabyl(survey$cluster_streamsprop) %>% 
  kable(caption = "Distribution of clusters from LCA on survey data")
```


We first interpret the classes based on the distribution of genres

```{r}
#| label: cluster from streams with proportion, interpretation
tb <- survey %>%
  select(hashed_id, cluster_streamsprop, matches("B_genres_\\d+")) %>%
  # consolidate genres
  pivot_longer(matches("B_genres_\\d+")) %>%
  mutate(name = genres_aliases[name],
         value = ifelse(value != "", 1, 0)) %>% 
  # remove empty genres
  filter(!is.na(name)) %>% 
  distinct() %>% 
  group_by(cluster_streamsprop, name) %>% 
  summarise(m=sum(value)/n()) %>% 
  pivot_wider(names_from = cluster_streamsprop, values_from = m) %>% 
  rename(genre = "name")
tb %>% 
  adorn_pct_formatting() %>% 
  kable(caption = "Profils of clusters")
```

```{r}
#| label: plot LPA streams prop cluster profiles
#| fig.height: 8
tb %>% 
  pivot_longer(-genre) %>% 
  ggplot(aes(value, genre, fill=genre)) +
    geom_col() +
    facet_wrap(~name)
```

```{r}
#| label: plot LPA streams prop cluster profiles most significant genres
#| fig.height: 8

sig_genres <- c("rock", "classical", "jazz", "frenchrap", "frenchsongs")
tb %>% 
  filter(genre %in% sig_genres) %>% 
  pivot_longer(-genre) %>% 
  ggplot(aes(value, genre, fill=genre)) +
    geom_col() +
    facet_wrap(~name)
```

We then look at how these classes are distributed in the population, 
across gender, age, and SES.

```{r}
#| label: cluster from streams with proportion, demographic data
cluster_dem <- survey %>% 
  select(cluster_streamsprop, age, gender, degree) %>% 
  filter(!is.na(cluster_streamsprop))
```


```{r}
#| label: cluster from streams with proportion, age
cluster_dem %>% 
  ggplot(aes(age, factor(cluster_streamsprop))) +
    geom_boxplot()
```

```{r}
#| label: cluster from streams with proportion, degree
cluster_dem %>% 
  filter(!is.na(degree)) %>% 
  ggplot(aes(factor(cluster_streamsprop), fill=degree)) +
    geom_bar(position="fill")
```

```{r}
#| label: cluster from streams with proportion, gender
cluster_dem %>% 
  filter(!is.na(gender)) %>% 
  ggplot(aes(factor(cluster_streamsprop), fill=gender)) +
    geom_bar(position="fill")
```

```{r}
#| label: cluster from streams with proportion, omnivorous
survey %>% 
  select(cluster_streamsprop, starts_with("omni"), mean_exo_pca, sd_exo_pca) %>% 
  pivot_longer(-cluster_streamsprop) %>% 
  mutate(name = recode_vars(name, "legitimacy")) %>% 
  filter(!is.na(value)) %>% 
  group_by(cluster_streamsprop, name) %>% 
  summarize(value = mean(value)) %>% 
  pivot_wider(names_from = name, values_from = value) %>% 
  kable()
```

## Comparison between survey and streaming

Do the clusters match?

```{r}
#| label: comparing the two LPA on streams
tabyl(survey, cluster_streamsprop, cluster_streams) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  kable(caption = "Relation between clusters from streams and from streams with proportions")
```

```{r}
#| label: comparing survey with streams
tabyl(survey, cluster_survey, cluster_streams) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  kable(caption = "Relation between cluster from survey and cluster from streams")
```


```{r}
#| label: comparing survey with streams prop
tabyl(survey, cluster_survey, cluster_streamsprop) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  kable(caption = "Relation between cluster from survey and cluster from streams with proportion")
```

## Legitimacy by cluster

Now, how does the legitimacy of artists from some genres vary along clusters?

We look at the distribution of legitimacy genre by genre, cluster by cluster.

TODO: RESCALE THESE BY GENRE (AT THE ARTIST LEVEL) FOR HOMOGENOUS X

```{r}
#| label: legitimacy of specific genres by cluster, survey
#| fig.height: 12
#| fig.width: 8
survey %>% 
  select(cluster_survey, starts_with("mean_exo_pca_")) %>% 
  pivot_longer(-cluster_survey) %>% 
  filter(!is.na(cluster_survey), !is.na(value)) %>% 
  mutate(name = str_remove(name, "mean_exo_pca_")) %>% 
  group_by(cluster_survey, name) %>% 
  summarize(m = mean(value), se = 1.96 * sd(value) / sqrt(n())) %>% 
  ggplot(aes(y=m, x=factor(cluster_survey), ymin = m-se, ymax=m+se)) +
    geom_point() +
    geom_linerange() +
    coord_flip() +
    facet_wrap(~name, scales = "free_x") +
    labs(x = "Mean exo. leg.", y = "Cluster", title = "Mean legitimacy of music from each genre")
```

```{r}
#| label: legitimacy of specific genres by cluster, streams
#| fig.height: 12
#| fig.width: 8
survey %>% 
  select(cluster_streams, starts_with("mean_exo_pca_")) %>% 
  pivot_longer(-cluster_streams) %>% 
  filter(!is.na(cluster_streams), !is.na(value)) %>% 
  mutate(name = str_remove(name, "mean_exo_pca_")) %>% 
  group_by(cluster_streams, name) %>% 
  summarize(m = mean(value), se = 1.96 * sd(value) / sqrt(n())) %>% 
  ggplot(aes(y=m, x=factor(cluster_streams), ymin = m-se, ymax=m+se)) +
    geom_point() +
    geom_pointrange() +
    coord_flip() +
    facet_wrap(~name, scales = "free_x") +
    labs(x = "Mean exo. leg.", y = "Cluster", title = "Mean legitimacy of music from each genre")

```

```{r}
#| label: legitimacy of specific genres by cluster, streams prop
#| fig.height: 12
#| fig.width: 8
survey %>% 
  select(cluster_streamsprop, starts_with("mean_exo_pca_")) %>% 
  pivot_longer(-cluster_streamsprop) %>% 
  filter(!is.na(cluster_streamsprop), !is.na(value)) %>% 
  mutate(name = str_remove(name, "mean_exo_pca_")) %>% 
  group_by(cluster_streamsprop, name) %>% 
  summarize(m = mean(value), se = 1.96 * sd(value) / sqrt(n())) %>% 
  ggplot(aes(y=m, x=factor(cluster_streamsprop), ymin = m-se, ymax=m+se)) +
    geom_point() +
    geom_pointrange() +
    coord_flip() +
    facet_wrap(~name, scales = "free_x") +
    labs(x = "Mean exo. leg.", y = "Cluster", title = "Mean legitimacy of music from each genre")

```

# Second strategy: looking at continuous metrics

```{r}
#| eval: false
library(GGally)
select(survey, starts_with("omni"), mean_exo_pca, sd_exo_pca) %>% 
  ggpairs()
```

## The link between diversity over legitimacy level and other measures of omnivorousness

```{r}
#| label: pairwise omnivorousness, exo vs. HHI streams
survey %>% 
  ggplot(aes(omni_stream_genres_hhi, sd_exo_pca)) +
    geom_point() +
  geom_smooth(method="lm") +
        labs(title = "The higher the diversity of genres actually played,\n the lower the diversity of exogenous legitimacy",
         x = "Diversity of genres played (HHI on streaming data)", y = "exo omnivorousness")
```

```{r}
#| label: pairwise omnivorousness, exo vs. richness listened
survey %>% 
  filter(!is.na(omni_survey_sum_genres_played)) %>% 
  ggplot(aes(factor(omni_survey_sum_genres_played), sd_exo_pca)) +
    geom_boxplot() +
    labs(title = "The more genres someone declares to listen, the lower the diversity over legitimacy levels",
         x = "# of genres declared played", y = "exo omnivorousness") +
    ylim(c(0, .4))
```

## Demographic correlates of HHI omnivorousness

Reminder: HHI omnivorousness is (1-herfindahl index) and indicates the
diversity of music streaming across genres (0 is low, 1 is high).

```{r}
#| label: HHI omnivorousness by gender and degree
select(survey, omni_stream_genres_hhi, gender, degree) %>% 
  filter(!is.na(gender), !is.na(degree)) %>% 
  pivot_longer(-omni_stream_genres_hhi) %>% 
  group_by(name, value) %>% 
  summarize(m = mean(omni_stream_genres_hhi, na.rm=TRUE),
            sd = sd(omni_stream_genres_hhi, na.rm=TRUE),
            n = n(),
            se = 1.96*sd/sqrt(n)) %>% 
  ggplot(aes(value, m, ymin = m-se, ymax=m+se)) +
    geom_point() +
    geom_pointrange() +
    facet_wrap(~name, scales = "free_x") +
    labs(title = "HHI omnivorousness by degree and gender", x="", y = "HHI omnivorousness")
```

```{r}
#| label: HHI omnivorousness by gender and degree, boxplot
select(survey, omni_stream_genres_hhi, gender, degree) %>% 
  filter(!is.na(gender), !is.na(degree)) %>% 
  pivot_longer(-omni_stream_genres_hhi) %>% 
  ggplot(aes(value, omni_stream_genres_hhi)) +
    geom_boxplot(notch = TRUE) +
    ylim(c(0, .25)) +
    facet_wrap(~name, scales = "free_x") +
    labs(title = "HHI omnivorousness by degree and gender", x="", y = "HHI omnivorousness")

```

```{r}
#| label: HHI omnivorousness by age
select(survey, omni_stream_genres_hhi, age) %>% 
  ggplot(aes(age, omni_stream_genres_hhi)) +
    geom_point() +
    geom_smooth(method="lm") +
    labs(x = "age", y="HHI omnivorousness",
         title = "The older one is, the higher the HHI omnivorouness")

```


## Demographic correlates of exo omnivorousness

```{r}
#| label: exo omnivorousness by gender and degree
select(survey, sd_exo_pca, gender, degree) %>% 
  filter(!is.na(gender), !is.na(degree)) %>% 
  pivot_longer(-sd_exo_pca) %>% 
  group_by(name, value) %>% 
  summarize(m = mean(sd_exo_pca, na.rm=TRUE),
            sd = sd(sd_exo_pca, na.rm=TRUE),
            n = n(),
            se = 1.96*sd/sqrt(n)) %>% 
  ggplot(aes(value, m, ymin = m-se, ymax=m+se)) +
    geom_point() +
    geom_pointrange() +
    facet_wrap(~name, scales = "free_x") +
    labs(title = "Exo omnivorousness by degree and gender", x="", y = "exo omnivorousness")
```

```{r}
#| label: exo omnivorousness by gender and degree, boxplot
select(survey, sd_exo_pca, gender, degree) %>% 
  filter(!is.na(gender), !is.na(degree)) %>% 
  pivot_longer(-sd_exo_pca) %>% 
  ggplot(aes(value, sd_exo_pca)) +
    geom_boxplot(notch = TRUE) +
    ylim(c(0, .25)) +
    facet_wrap(~name, scales = "free_x") +
    labs(title = "Exo omnivorousness by degree and gender", x="", y = "exo omnivorousness")

```

```{r}
#| label: exo omnivorousness by age
select(survey, sd_exo_pca, age) %>% 
  ggplot(aes(age, sd_exo_pca)) +
    geom_point() +
    geom_smooth(method="lm") +
    labs(x = "age", y="exo omnivorousness",
         title = "The older one is, the lower the diversity of exogenous legitimacy")

```

# Misc

Definition of cultural holes omnivorousness

$$EO_i = \sum_{j \in N(i)}(a_{ij} - (\frac{1}{OV_i-1}\sum_{k \in N(i)}^{k \neq j}{o_{jk}}))$$

